---
title: "R_Tables"
author: "M.O."
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Use AI to Create a Simulated Dataset

**Instructions:** Please generate R code to create a dataset of 4000 people and three variables. Variable 1 is called 'agegroup' and it contains four groups called '45-54 Years', '55-64 Years', '65-74 years', and '75+ Years'. 1000 people are in each category. Variable 2 is called 'depr' and it ranges from 0-30, the mean score is 7 and the standard deviation is 10, but everyone in the 75th percentile should be in the '75+ Years' group. Variable 3 is called 'fss' and ranges from 0-100, with a mean of 85 and an SD of 20, but everyone in the 25th percentile should be in the '75+ Years' group.

AI response: Here’s an R script that creates a dataset of 4000 people with the specified characteristics:

```{r}
# Set seed for reproducibility
set.seed(123)

# Generate 'agegroup' variable
agegroup <- rep(c('45-54 Years', '55-64 Years', '65-74 Years', '75+ Years'), each = 1000)

# Generate 'depr' variable with constraints
depr <- c(
  rnorm(1000, mean = 7, sd = 10),                # 45-54 Years
  rnorm(1000, mean = 7, sd = 10),                # 55-64 Years
  rnorm(1000, mean = 7, sd = 10),                # 65-74 Years
  rnorm(1000, mean = 7, sd = 10)                 # 75+ Years
)

# Ensure everyone in the 75th percentile of 'depr' is in the '75+ Years' group
depr_75th <- quantile(depr, 0.75)
agegroup[depr > depr_75th] <- '75+ Years'

# Generate 'fss' variable with constraints
fss <- c(
  rnorm(1000, mean = 85, sd = 20),               # 45-54 Years
  rnorm(1000, mean = 85, sd = 20),               # 55-64 Years
  rnorm(1000, mean = 85, sd = 20),               # 65-74 Years
  rnorm(1000, mean = 85, sd = 20)                # 75+ Years
)

# Ensure everyone in the 25th percentile of 'fss' is in the '75+ Years' group
fss_25th <- quantile(fss, 0.25)
agegroup[fss < fss_25th] <- '75+ Years'

# Create the final dataset
data <- data.frame(
  agegroup = agegroup,
  depr = depr,
  fss = fss
)

# View the first few rows of the dataset
# head(data)
```

## Transform agegroup into a factor variable and order the response categories in ascending order

```{r}
data$agegroup <- factor(data$agegroup, levels = c("45-54 Years", "55-64 Years", "65-74 Years", "75+ Years"))
```

---

# Load the Libraries of Interest

```{r}
library(gtsummary)
library(broom)
library(huxtable)
library(openxlsx)
```

---

# Descriptively Summarize the Data

gtsummary package: <https://www.danieldsjoberg.com/gtsummary/index.html>

*include =* specifies the variables we want in the table - leave out this command if you want all variables to show in the table.

*label = list* assigns meaningful variable names for the table.

Default is to print median and Q1, Q3 statistics in the the table, so *statistic = list(all_continuous() ~ "{mean} ({sd})")* changes the output to mean and SD.

```{r}
theme_gtsummary_journal(journal = "jama") # These first two lines of code format the table according to JAMA guidelines.
theme_gtsummary_compact()

table1 <-
  data |> tbl_summary(include = c(agegroup, depr, fss), 
                      statistic = list(all_continuous() ~ "{mean} ({sd})"),
                      label = list(agegroup ~ "Age Group", depr ~ "Depression Symptoms", fss ~ "Functional Social Support"))

# View the table in R by typing 'table1' in the R console.
```
## Output the table as a CSV file for formatting and placement into a Word file

```{r}
gtsummary::as_hux_xlsx(table1, file="table1.csv") # Specify the full file path if you are not working in a project
```

---

# Run a Regression Analysis & Get Tabular Output

## Run the regression

Regress depr on fss and agegroup in a multiple linear regression model.

Here is a good tutorial on running linear regression in R: <https://www.scribbr.com/statistics/linear-regression-in-r/>

```{r}
model1 <- lm(depr ~ fss + agegroup, data = data)

summary(model1)
confint(model1)

# The default output is not pleasant on the eye.
```
## Get tabular and forest plot output

*label = list* assigns meaningful variable names for the table.

```{r}
table2 <- 
  model1 |> tbl_regression(label = list(fss ~ "Functional Social Support", agegroup ~ "Age Group"))

# View the table in R by typing 'table2' in the R console.

forest_plot <- 
  model1 |> tbl_regression(label = list(fss ~ "Functional Social Support", agegroup ~ "Age Group")) |> plot()

# View the forest plot in R by typing 'forest_plot' in the R console.
```

## Output the table as a CSV file for formatting and placement into a Word file

```{r}
gtsummary::as_hux_xlsx(table2, file="table2.csv") # Specify the full file path if you are not working in a project
```